\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-30 OaklandSubmission/ex.tex
super minor
minor

\subsubsection*{Source language}
Consider the example $w^Tx >b$ from Section~\ref{sec:intro}, where
$w$ and $b$ constitute the server's input (a classifier) and $x$ is
the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
for this example. The code first reads the server's (client's) input using
 \ls{input1} (\ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
\ls{x}. Finally, the code outputs the result of comparing \ls{acc}
with \ls{b} only to the client using \ls{output2}.

@@

The compiler splits the input
program into \emph{public} and \emph{secret} components. The public
components translate into regular C++ code, while the secret
components translate into API calls into our crypto back-end
(ABY). E.g., for the code in Figure~\ref{fig:ex-sml}, the \tool
compiler realizes that the array index \ls{i} in the dot product loop
is public, and hence the array accesses need not be compiled
obliviously. Therefore, it compiles the \ls{for}~loop into a C++
\ls{for}~loop that will be executed in-clear
(line~\ref{line:dotproductloop}).

Within the secret components, the \tool compiler is ``cryptographic
cost-aware'', and appropriately picks either arithmetic or boolean
circuit representations for different sub-componenets. For example,
the compiler realizes that the dot product computation is more
efficient in the arithmetic representation, and therefore it builds
the corresponding circuit using the arithmetic circuit builder
\ls{acirc} (lines~\ref{line:dotmulgate} and~\ref{line:dotaddgate}). On
the other hand, since the comparison with \ls{b}, and the conditional
expression computation are more efficient in the boolean
representation, the \tool compiler uses the Yao circuit builder
\ls{ycirc} to build the corresponding circuits
(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{As
  we mention in
  Section~\ref{sec:impl}, in our usage of ABY, boolean is synonymous
  with Yao.}.

Importantly, conversions between the arithmetic and boolean
representations require share-conversions. The \tool compiler also
instruments these conversion gates accordingly. For example, in
lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
\ls{a_acc} and \ls{a_b} to boolean representation, before they are
input to the comparison and multiplexer circuits.

@@

To summarize, \tool raises the level of abstraction for the
programmer, and generates efficient crypto protocols automatically,
while its metatheory provides strong correctness and security
guarantees. We provide a formal description of the \tool language next.

%% Consider a  cloud service provider Bob that wants to provide a service to diagnose
%% whether a patient has breast cancer or not. Bob trains a machine learning classification model
%% that results in a vector $w$ (say of length 30) and a scalar $b$.
%% This model is the intellectual property of Bob and he  wants to keep it secret.
%% Given a patient's medical report, in the form of a vector $x$ (also of length 30),
%% the classifer predicts that the patient has breast cancer if $w^Tx>b$.
%% However, for this task Bob needs access to $x$, which is private data of a customer.

>>>>

\subsubsection*{Source language}
Consider the example $w^Tx >b$ from Section~\ref{sec:intro}, where
$w$ and $b$ constitute the server's input (a classifier) and $x$ is
the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
for this example. The code first reads the server's (resp. client's) input using
 \ls{input1} (resp. \ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
\ls{x}. Finally, the code outputs the result of comparing \ls{acc}
with \ls{b} only to the client using \ls{output2}.

@@

The compiler splits the input
program into \emph{public} and \emph{secret} components. The public
components translate into regular C++ code, while the secret
components translate into API calls into our crypto back-end
(ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
compiler realizes that the array index \ls{i} in the dot product loop
is public, and hence the access locations need not be hidden.
Therefore, it compiles the \ls{for}~loop into a C++
\ls{for}~loop that will be executed in-clear
(line~\ref{line:dotproductloop}).

Within the secret components, the \tool compiler is ``cryptographic
cost-aware'', and appropriately picks either arithmetic or boolean
circuit representations for different sub-componenets. For example,
the compiler realizes that the dot product computation is more
efficient in the arithmetic representation, and therefore it builds
the corresponding circuit using the arithmetic circuit builder
\ls{acirc} (lines~\ref{line:dotmulgate} and~\ref{line:dotaddgate}). On
the other hand, since the comparison with \ls{b}, and the conditional
expression computation are more efficient in the boolean
representation, the \tool compiler uses the Yao circuit builder
\ls{ycirc} to build the corresponding circuits
(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{In our usage of ABY, boolean is synonymous
  with Yao (Section~\ref{sec:impl}).}.

Using both arithmetic and boolean representations requires conversions between them. The \tool compiler also
instruments these conversion gates accordingly. For example, in
lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
\ls{a_acc} and \ls{a_b} to boolean representation, before they are
input to the comparison and multiplexer circuits.

@@

To summarize, \tool raises the level of abstraction for the
programmer, and generates efficient crypto protocols automatically,
while its metatheory provides strong correctness and security
guarantees. 
%% Consider a  cloud service provider Bob that wants to provide a service to diagnose
%% whether a patient has breast cancer or not. Bob trains a machine learning classification model
%% that results in a vector $w$ (say of length 30) and a scalar $b$.
%% This model is the intellectual property of Bob and he  wants to keep it secret.
%% Given a patient's medical report, in the form of a vector $x$ (also of length 30),
%% the classifer predicts that the patient has breast cancer if $w^Tx>b$.
%% However, for this task Bob needs access to $x$, which is private data of a customer.

### 2017-10-30 OaklandSubmission/intro.tex
minor

%% Today it is hard for the developers to program secure applications
%% using cryptographic techniques. Typical developers lack a deep
%% understanding of the cryptographic protocols, and therefore cannot be
%% expected to use them correctly and efficiently on their own.
%% Ideally, a developer would declare the functionality in a general
%% purpose, high-level programming language and a
%% tool, e.g. a compiler, would generate an efficient protocol that
%% implements the functionality securely, while hiding the cryptography
%% behind-the-scenes. This paper presents such a framework for Secure
%% Two-party Computation (\mpc).

%Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs. It is critical to address this fundamental  representation gap to make cryptography accessible to non-crypto specialist developers. In this work, we address this gap for the case of secure multi-party computation.

%Most cryptographic tools are very hard to use non-crypto specialist developers. The task becomes even more tedious and error-prone for the case of secure multiparty computation since different protocol implementations require the function to be expressed in different representations such as Boolean or Arithmetic circuits. To make cryptography accessible to non-crypto specialist developers, we need to design systems that do crypto automatically and satisfy the properties of ...

Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is a powerful
cryptographic technique that allows two mutually distrusting parties
to compute a publicly known joint function of their secret inputs in a way that both
the parties learn nothing about the inputs of each other beyond what
is revealed by their (possibly different) outputs. For example, \mpc
can be used for {\em secure
prediction}~(\cite{shafindss,wu,barni,minionn,secureml}),
where one party (the server) holds a proprietary classifier to predict
a label (e.g. a disease, genomics, or spam detection), and the other
party (the client) holds a private input that it wants to run the
classifier on. Using \mpc guarantees that the server learns nothing
about the client's input or output, and that the client learns nothing
about the server classifier, beyond what is revelaed by the output
label.

Since its introduction in 1980s, there has been a long
sequence of work that has transformed 2PC from a mere theoretical tool to
something practically efficient and usable.
Our goal is to build a framework that enables programmers, who are not
experts in cryptography, to easily program efficient and scalable \mpc
protocols.  To understand the
state-of-the-art, let us consider an example underlying many
secure prediction algorithms. Say Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
the server classifier, and $x$ is the client's input vector. Further,
$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
inner product of  $w^{T}$ and $x$. Alice has following options.

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation.}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs devoid of any
cryptographic details. The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous works~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
These papers (and others) cite the inefficiency of generic \mpc as the major motivation
behind the design of specialized protocols. Using \tool, we empirically  demonstrate
the surprising fact that generic \mpc implementations are much more efficient than what they were believed to be. 
 %(up to a factor of 19x in our
%experiments, depending on the functionality) \nc{I added this but I
%  guess we are saying it many times, but might be nice to highlight
%  perf early no?} \aseem{My rationale was to save the numbers for
%  Evaluation summary and remain high-level here.} 
Below we describe
the salient features of \tool.

@@

\subsubsection*{Formal guarantees} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for a well-typed source program, the two semantics
successfully terminate
(e.g. there are no array index out-of-bounds errors), with identical
observable outputs. For
the security theorem, we formally reduce the security
of our compiler to the security of the  \mpc back-end
 (by also invoking our correctness theorem). We also prove a formal security theorem against semi-honest (or, ``honest but curious") adversaries for our partitioning
scheme (Section~\ref{sec:ld} and Section~\ref{sec:pipe}).

>>>>

 Today it is hard for the developers to program secure applications
 using cryptographic techniques. Typical developers lack a deep
 understanding of the cryptographic protocols, and therefore cannot be
 expected to use them correctly and efficiently on their own.
 Ideally, a developer would declare the functionality in a general
 purpose, high-level programming language and a
 tool, e.g. a compiler, would generate an efficient protocol that
 implements the functionality securely, while hiding the cryptography
 behind-the-scenes.

 This paper presents such a framework for Secure
 Two-party Computation (\mpc),
%
%Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs. It is critical to address this fundamental  representation gap to make cryptography accessible to non-crypto specialist developers. In this work, we address this gap for the case of secure multi-party computation.
%
%Most cryptographic tools are very hard to use non-crypto specialist developers. The task becomes even more tedious and error-prone for the case of secure multiparty computation since different protocol implementations require the function to be expressed in different representations such as Boolean or Arithmetic circuits. To make cryptography accessible to non-crypto specialist developers, we need to design systems that do crypto automatically and satisfy the properties of ... 
%
%Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is
a powerful
cryptographic technique that allows two mutually distrusting parties
to compute a publicly known joint function of their secret inputs in a way that both
the parties learn nothing about the inputs of each other beyond what
is revealed by their (possibly different) outputs. For example, \mpc
can be used for {\em secure
prediction}~(\cite{shafindss,wu,barni,minionn,secureml}),
where one party (the server) holds a proprietary classifier to predict
a label (e.g. a disease, genomics, or spam detection), and the other
party (the client) holds a private input that it wants to run the
classifier on. Using \mpc guarantees that the server learns nothing
about the client's input or output, and that the client learns nothing
about the server classifier, beyond what is revelaed by the output
label.

%Since its introduction in 1980s, there has been a long
%sequence of work that has transformed 2PC from a mere theoretical tool to
%something practically efficient and usable.
%Our goal is to build a framework that enables programmers, who are not
%experts in cryptography, to easily program efficient and scalable \mpc
%protocols. 
To understand the
state-of-the-art, let us consider an example underlying many
secure prediction algorithms. Say Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
the server classifier, and $x$ is the client's input vector. Further,
$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
inner product of  $w^{T}$ and $x$. Alice has following options.

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation.}, the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs devoid of any
cryptographic details. The generated protocols use combinations of
arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous works~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
These papers (and others) cite the inefficiency of generic \mpc as the major motivation
behind the design of specialized protocols. Using \tool, we empirically  demonstrate
the surprising fact that generic \mpc implementations are much more efficient than what they were believed to be. 
 %(up to a factor of 19x in our
%experiments, depending on the functionality) \nc{I added this but I
%  guess we are saying it many times, but might be nice to highlight
%  perf early no?} \aseem{My rationale was to save the numbers for
%  Evaluation summary and remain high-level here.} 
Below we describe
the salient features of \tool.

@@

\subsubsection*{Formal guarantees} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for a well-typed source program, the two semantics
successfully terminate
(e.g. there are no array index out-of-bounds errors), with identical
observable outputs. For
the security theorem, we formally reduce the security
of the compiler output, against semi-honest (or, ``honest but curious") adversaries,
 to the security of the  \mpc back-end. 
 We also prove a formal security theorem against semi-honest  adversaries for our partitioning
scheme (Section~\ref{sec:ld} and Section~\ref{sec:pipe}).

### 2017-10-30 OaklandSubmission/ld.tex
minor

\subsubsection*{Source to intermediate compilation judgments} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme of our implementation.

@@

As mentioned earlier, the intermediate language models the code such as in \figureref{ex-aby} output by our compiler. 
Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later. The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure. 
Crucially, this phase of the semantics does not have access to the secret inputs.
Below, we first provide the language for the circuits used followed by our evaluation rules.

>>>>

\subsubsection*{Source to intermediate compilation} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme of our implementation.

@@

As mentioned earlier, the intermediate language models the code such as in \figureref{ex-aby} output by our compiler. 
Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed runtime later. The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure. 
Crucially, this phase of the semantics does not have access to the secret inputs.
Below, we first provide the language for the circuits followed by the evaluation rules.

### 2017-10-30 OaklandSubmission/pipe.tex
minor

To overcome this problem, we define a sequence of new programs
$\prog'_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
= \bot$. Without
loss of generality, let all $\state_i$ be values in some additive ring
$(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$,
i.e., the additive ring of integers modulo $2^{64}$).
Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from
the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all
$r_i$ values are generated by a pseudorandom function). Let $\prog'_1$
be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
(and empty state $\state_0$), and runs $\prog_1$ (as defined above) to
compute $\state_1$ and then outputs $o_1 = \state_1 + r_1$ {\em only
  to} Bob\footnote{While the description of our protocol here assumes
  that
  the underlying backend supports only one party receiving output,
  this is only a simplifying assumption, and we can easily modify our
  protocol in the case where both parties must receive the same
  output. To do so, we modify $\prog'_1$ to output $o_1 = s_1+r_1+t_1$ to
  both parties (where $t_1$ is random and chosen by Bob). We can then
  appropriately modify the remaining steps as well.}. Alice's output
from $\prog'_1$ is $r_1$. Next, every $\prog'_i$ ($2\leq i\leq k-1$)
takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob,
runs $\prog_i$
on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes
subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
only to Bob. Again, Alice's output from $\prog_i$ is $r_i$. The last
program $\prog'_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
parties.

>>>>

To overcome this problem, we define a sequence of new programs
$\prog'_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
= \bot$. Without
loss of generality, let all $\state_i$ be values in some additive ring
$(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$,
i.e., the additive ring of integers modulo $2^{64}$).
Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from
the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all
$r_i$ values are generated by a pseudorandom function). Let $\prog'_1$
be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
(and empty state $\state_0$), and runs $\prog_1$ (as defined above) to
compute $\state_1$ and then outputs $o_1 = \state_1 + r_1$ {\em only
  to} Bob\footnote{While the description of our protocol here assumes
  that
  the underlying backend supports only one party receiving output,
  this is only a simplifying assumption, and we can easily modify our
  protocol in the case where both parties must receive the same
  output. To do so, we modify $\prog'_1$ to output $o_1 = s_1+r_1+t_1$ to
  both parties (where $t_1$ is random and chosen by Bob). We can then
  appropriately modify the remaining steps as well.}. Alice's output
from $\prog'_1$ is $r_1$. Next, every $\prog'_i$ ($2\leq i\leq k-1$)
takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob,
runs $\prog_i$
on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes
subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
only to Bob and $r_i$ only to Alice. The last
program $\prog'_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
parties.
